Modeling the Java Card API (or Interface?) in B

### 2012-03-29 trunk/papers/B2012-work/javacardAPIpaper.tex

A Java Card applet is a class that inherits the
\emph{javacard.framewok.Applet} class of the Java Card API and is
implemented upon the Java Card subset of Java.  During the applet
conversion for card installation, a verification phase is performed to
check conformance of the classes with Java Card restrictions.

The current usual Java Card specification (2.2.x) allows two kinds of
applets.  The older, and most commonly used, kind of applet manipulates
directly the APDU protocol packages while the newer one abstracts from
the lower level protocol using Remote Method Invocation (RMI). In this
paper we will call the lower level applet \emph{APDU applet} and the
higher level one, \emph{RMI applet}. RMI introduces a layer of
abstraction above the APDU protocol, and, due to this fact, it is usually
less efficient than APDU applets.

@@

Among the different components of the Java Card platform, the Java
Card API is the one without which no refinement into Java Card code
could be done.  Suppose, for instance, that we had the B abstract
machine of Fig.~\ref{lis:buscard}, which models a bus card with a
variable that keeps track of the current credit the user has (\texttt{balance}) and an operation to add credits to this
balance (\texttt{add}). This simple example provides enough evidence of the
importance of having a B model of the Java Card API.  As all data in
it is implementable in C, for instance, and it is a deterministic
specification, B tools can directly generate code for it in C, and the
resulting code would be very similar to the original B machine.

@@

We can identify two main kinds of transformation that had to be carried out in order to obtain the code  of operation \texttt{add} in figs. \ref{lis:addjavacard} and \ref{lis:addimplementation} from the abstract specification in fig. \ref{lis:buscard}:

@@

\item the inclusion of Java Card specific aspects, such as the change of interface of \texttt{add} due to the need of passing data via the APDU protocol and the manipulation of the APDU buffer by methods offered the Java Card API.

@@

The transition into a defensive style of programming is part of the BSmart method, and implemented by the BSmart Full Function component. This component generates a B refinement of the original specification where pre-conditions are transformed into conditionals (IF-THEN-ELSE-END constructs), with the indication of the need to throw the corresponding exception when the specified conditions are not met by input data and states.
The resulting B model can then be formally verified as usual in the B method.

The change of interface is also taken care of by the BSmart method, so that the development can be done without need to extend the B method. More details on this [[[???? (we haven't published this part but there is the technical report )]]].

@@

One delicate issue to be dealt with both by the user of the API specification and the tools designed to work with it, is the use of overloaded methods in an application development. When overloaded operations are used in an application development, the user must explicitly choose which version of the overloaded operation is to be used. Then, verifications will be carried out considering this specific version of the operation. When code is generated, either the code generator will remove the name extension automatically, if designed for it, or the user will have to edit the generated code manually. In any case, it is important that the actual method called in the Java Card code is the one with which verifications have been carried out.   In case of overriding, in particular, it will the user's responsibility to choose in B the one which is on scope at this specific point in Java, unless a extra verification step is included in the code generator that checks that the operation used in B is the one in scope in Java and registers an error in case it is not.

>>>>

A Java Card applet is a class that inherits the
\emph{javacard.framewok.Applet} class of the Java Card API and is
implemented upon the Java Card subset of Java.  During the applet
conversion for card installation, a verification phase is performed to
check conformance of the classes to Java Card restrictions.

The current usual Java Card specification (2.2.x) allows two kinds of
applets.  The older, and most commonly used, kind of applet
manipulates directly the APDU protocol packages while the newer one
abstracts from the lower level protocol using Remote Method Invocation
(RMI). In this paper we will call the lower level applet \emph{APDU
  applet} and the higher level one, \emph{RMI applet}. RMI introduces
a layer of abstraction above the APDU protocol, and hence is usually
less efficient than APDU applets.

@@

Among the different components of the Java Card platform, the Java
Card API is the one without which no refinement into Java Card code
could be done.  Suppose, for instance, that we had the B abstract
machine of Fig.~\ref{lis:buscard}, which models a bus card with a
variable that keeps track of the current credit the user has
(\texttt{balance}) and an operation to add credits to this balance
(\texttt{add}). This simple example provides enough evidence of the
importance of having a B model of the Java Card API.  As all data in
it is implementable in C, for instance, and as it is a deterministic
specification, B tools can directly generate code for it in C, and the
resulting code would be very similar to the original B machine.

@@

We can identify two main kinds of transformation that have to be
carried out in order to obtain the code of operation \texttt{add} in
figs.~\ref{lis:addjavacard} and~\ref{lis:addimplementation} from the
abstract specification in fig.~\ref{lis:buscard}:

@@

\item the inclusion of Java Card specific aspects, such as changing
  the signature of \texttt{add} so that data is transmitted via the
  APDU protocol and the APDU buffer is manipulated by methods offered
  the Java Card API.

@@

The transition to a defensive style of programming is part of the
BSmart method, and implemented by the BSmart Full Function
component. This component generates a B refinement of the original
specification where pre-conditions are transformed into conditionals
(IF-THEN-ELSE-END constructs), with the indication of the need to
throw the corresponding exception when the specified conditions are
not met by input data and states.  The resulting B model can then be
formally verified as usual in the B method.

The change of interface is also taken care of by the BSmart method, so
that the development can be done without need to extend the B
method, as detailed in~\cite{Gomes10}.

@@

One delicate issue to be dealt with both by the user of the API
specification and the tools designed to work with it, is the use of
overloaded methods in an application development. When overloaded
operations are used in an application development, the user must
explicitly choose which version of the overloaded operation is to be
used. Then, verifications will be carried out considering this
specific version of the operation. When code is generated, either the
code generator will remove the name extension automatically, if
designed for it, or the user will have to edit the generated code
manually. In any case, it is important that the actual method called
in the Java Card code is the one with which verifications have been
carried out.  In case of overriding, in particular, it will be either
the user's responsibility to choose the B one which is in scope at
this specific point in Java, or an extra verification step will be
included in the code generator to check that the operation used in
B is the one in scope in Java and registers an error in case it is
not.
